Generalized spaces for constructive algebra

### 2020-04-05 paper-herrsching2019/main.tex
Fixed typography of exercises

\documentclass{ws-rv9x6}
\usepackage[onethmnum]{ws-rv-thm}
\usepackage{ws-rv-van}
\usepackage{mathtools,stmaryrd,xspace,bussproofs}

\newlength{\exerciseskip}
\setlength{\exerciseskip}{1.5em}
\newcounter{exercisenr}
\newenvironment{exercise}[1]{
  \refstepcounter{exercisenr}
  \noindent\textbf{Exercise \theexercisenr{}.} \emph{#1} \smallskip\par\noindent%
}{\par\vspace{\exerciseskip}}

\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\usepackage{tikz}
\usetikzlibrary{calc}
\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red!80, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

@@

\begin{exercise}{The topological space of points of a locale}%
\label{ex:canonical-topology}%
Devise a canonical topology on the set of points of a locale.
\end{exercise}

\begin{exercise}{On the embedding of sober topological spaces into locales}%
Verify the statements in Remark~\ref{rem:sober-embed}.
\end{exercise}

\begin{exercise}{Hausdorff spaces are sober}%
\label{ex:hausdorff-sober}%
A \emph{meet-irreducible open subset} of a topological space~$Y$ is an
open subset~$W \subseteq Y$ such that~$(U \cap V \subseteq W) \Rightarrow (U
\subseteq W) \vee (V \subseteq W)$ for all open subsets~$U,V \subseteq Y$ and
such that~$Y \not\subseteq W$.
\begin{enumerate}
\item Assuming the law of excluded middle, show that the completely prime
filters of a topological space are in canonical one-to-one correspondence with
the meet-irreducible open subsets.

@@

\item Assuming the law of excluded middle, show that Hausdorff topological
spaces validate the condition stated in part~(b) and hence are sober.
\end{enumerate}
\end{exercise}

@@

\begin{exercise}{On the covariant approach}%
Let~$A$ be a frame.
\begin{enumerate}
\item Show that~$A$ contains arbitrary (set-indexed) meets (though they are not
required to be preserved by frame homomorphisms), by the impredicative
construction
\[ \bigwedge_{i \in I} U_i = \bigvee\{ \text{$V \preceq U_i$ for all~$i \in I$}
\}. \]
\item What are the arbitrary meets in the frame of open subsets of a
topological space?
\item Show that for any frame homomorphism~$\alpha : A \to A'$ there exists a
monotone map~$\beta : A' \to A$ such that~\emph{$\beta$ is right adjoint
to~$\alpha$}, that is such that
\[ \alpha(U) \preceq V \quad\text{iff}\quad U \preceq \beta(V) \]
for all~$U \in A$ and~$V \in A'$.
\end{enumerate}
{\scriptsize\emph{Note.} Based on the observation in part~(c), a
\emph{covariant approach} to locales can be developed, in which locale maps~$X
\to X'$ are defined to be certain kinds of maps~$\O(X) \to
\O(X')$~\cite{picado-pultr:covariant}.\par}
\end{exercise}

@@

\begin{exercise}{Characterizing spatial locales}%
Let~$X$ be a locale. Show that the following conditions are equivalent.
\begin{enumerate}
\item For any opens~$U, V \in \O(X)$: If~$x \inplus U \Rightarrow y \inplus V$
for all points~$x$ of~$X$, then~$U \preceq V$.
\item For any opens~$U, V \in \O(X)$: If~$x \inplus U \Leftrightarrow y \inplus V$
for all points~$x$ of~$X$, then~$U = V$.
\item There is a topological space~$Y$ and a surjective morphism~$L(Y) \to X$
of locales. (A morphism of locales is \emph{surjective} iff its underlying
frame homomorphism is injective.)
\item The canonical morphism~$L(\Pt(X)) \to X$ is an isomorphism.
\end{enumerate}
\end{exercise}

\begin{exercise}{The frame of radical ideals}%
\label{ex:frame-of-radical-ideals}%
Let~$A$ be a ring. We order the set~$\Idl(A)$ of ideals of~$A$ by inclusion.
(See Section~\ref{sect:algebraic-preliminaries} for some algebraic
preliminaries.)
\begin{enumerate}
\item Verify that~$\Idl(A)$ has arbitrary meets and joins, given by
intersection and sum of ideals, but that it is in general not a frame.

{\scriptsize\emph{Hint.} The distributive law fails in the polynomial ring~$\QQ[X,Y]$.\par}
\item Show that the sub-poset~$\Rad(A)$ of radical ideals is a frame, with
the same finite meets but with joins given by the radical of the join computed
in~$\Idl(A)$.
\item\label{item:join-top} Let~$\varphi$ be a truth value. Let~$\{\top\,|\,\varphi\}$ be the
subsingleton subset of~$\Rad(A)$ which contains the top element~$\sqrt{(1)}$
iff~$\varphi$. Verify that
\[ \bigvee\{\top\,|\,\varphi\} = \{ f \in A \,|\, \text{$f$ is nilpotent
or~$\varphi$} \} \]
in~$\Rad(A)$.
\end{enumerate}
\end{exercise}

\begin{exercise}{Coproducts of locales}%
\begin{enumerate}
\item Show that the category of frames has small products.
\item Deduce that the category of locales has small coproducts.
\item Let~$M$ be a set and endow it with the discrete topology. Show
that~$L(M)$ is the coproduct~$\coprod_{x \in M} \pt$.
\end{enumerate}
{\scriptsize\emph{Note.} The category of locales also has small products;
Exercise~\ref{ex:products-of-locales} is devoted to this fact.\par}
\end{exercise}

@@

\begin{exercise}{Localic maps into discrete spaces}%
Let~$X$ be a locale. Let~$A$ be a set and endow it with the discrete topology. Show
that morphisms~$X \to L(A)$ are in canonical correspondence with
families~$(U_a)_{a \in A}$ of opens of~$X$ such that~$\top = \bigvee_{a \in A}
U_a$ and~$U_a \wedge U_b \preceq \bigvee\{ \top \,|\, a = b \}$.
\end{exercise}

@@

\begin{exercise}{Discrete locales as classifying locales}%
\begin{enumerate}
\item Find a theory which is classified by the two-point locale (the locale
associated to the discrete two-point topological space).
\item Let~$M$ be a set and endow it with the discrete topology. Find a theory
which is classified by~$L(M)$.
\end{enumerate}
\end{exercise}

@@

\begin{exercise}{More on exotic locales}%
\label{ex:exotic-locales}%
\begin{enumerate}
\item Define the locale of injections~$\RR \hookrightarrow \NN$.
\item Define the locale of partial surjections~$\NN \twoheadrightarrow \NN^\NN$.

{\scriptsize\emph{Note.} In Russian constructivism, as for instance reigning in
the ef{}fective topos~\cite{hyland:effective-topos,phoa:effective,bauer:c2c}, the Baire space~$\NN^\NN$ is subcountable.
However, in this setting any function~$\NN \to \NN$ is computable.
The universe of sheaves over this locale was the first example of a setting in
which the Baire space is subcountable and still any external number-theoretic
function can be imported~\cite[Section~4.3]{scedrov:forcing}.\par}
\item Define the locale of ``open covers of~$[0,1]$ by intervals with rational
endpoints such that the union of any finite number of those intervals has total
Lebesgue measure less than~$\frac{1}{2}$''.

{\scriptsize\emph{Note.} In the universe of sheaves over this locale, the set
of unit interval Dedekind real numbers fails to be (Heine--Borel) compact and
the localic real line fails to be
spatial~\cite[Example~D4.7.13]{johnstone:elephant}.\par}
\item An open~$U$ of a locale~$X$ is \emph{complemented} iff there is an
open~$V$ such that~$U \vee V = \top$ and~$U \wedge V = \bot$. A locale is
\emph{zero-dimensional} iff any open is a join of complemented opens. Assuming
that any two reals are equal or not, show that the locale of surjections~$\NN
\twoheadrightarrow \RR$ is zero-dimensional.
\end{enumerate}
\end{exercise}

\begin{exercise}{Models of theories as points of the classifying locales}%
Let~$\TT$ be a propositional geometric theory. Verify that the points
of~$L(\TT)$ are in canonical bijection with the set-based models of~$\TT$.
\end{exercise}

@@

\begin{exercise}{Spatiality of the localic reals}%
\label{ex:spatiality-reals}%
\newcommand{\fa}{\ensuremath{(\!\cdot}}%
\newcommand{\fb}{\ensuremath{\cdot\!)}}%
Let~$\RR$ be the localic real line, defined as the classifying locale of the
theory~$\TT$ of Dedekind cuts as as introduced on
page~\pageref{par:dedekind-cuts}. For rational numbers~$x,y$, let~``$\fa x,y
\fb$'' denote the formula~$\alpha_x \wedge \beta_y$. Let~$(x,y)$ denote the
open interval~$\{ \xi \in \Pt(\RR) \,|\, x < \xi < y \}$ in the set of Dedekind reals.
\begin{enumerate}
\item Extend the notation to allow for~$x,y=\pm\infty$.
\item Show that the canonical topology on~$\Pt(\RR)$, defined in
Exercise~\ref{ex:canonical-topology}, coincides with the usual Euclidean
topology.
\item Show that any open of~$\RR$ can be written in the form~$\bigvee_i
[\fa x_i,y_i \fb]$ with~$x_i,y_i \in \QQ \cup \{\pm\infty\}$.
\item Let rational numbers be given such that~$(x,y) \subseteq \bigcup_i
(x_i,y_i)$ in~$\Pt(\RR)$. Assume that this open covering has a Lebesgue number.
Show that~$\TT$ proves~$\fa x,y \fb \vdash \bigvee_i \fa x_i,y_i \fb$.
\item Assume that closed intervals in the set of Dedekind reals are
(Heine--Borel) compact. Then redo part~(d), but without the assumption on the
existence of a Lebesgue number.
\item Under the same assumption as in part~(e), show that the localic real line
is spatial.
\end{enumerate}
{\scriptsize\emph{Note.} Exercise~\ref{ex:exotic-locales} lists settings in
which the assumption in part~(e) is not satisfied.\par}
\end{exercise}

\begin{exercise}{The localic Galois group of a field extension}%
Let~$L|k$ be a field extension. The \emph{localic Galois group}~$\Gal(L|k)$
is the classifying locale of the theory of automorphisms~$L \to L$ which
fix~$k$, as described on page~\pageref{par:theory-automorphisms}.
\begin{enumerate}
\item Explain how to obtain a morphism~$\Gal(L|k) \to \Gal(L|k)$ which on the
level of points should send an automorphism~$\sigma$ to its inverse~$\sigma^{-1}$.
\item Explain how to obtain a morphism~$\Gal(L|k) \times \Gal(L|k) \to \Gal(L|k)$ which on the
level of points should send a pair~$\langle \sigma, \tau \rangle$ to~$\sigma \circ \tau$.
\item Assume from now on that~$L$ is a \emph{geometric field} in the sense that any
element is zero or invertible. Show that the axiom expressing injectivity can
be dropped without changing the set of provable sequents.
\item Let~$f \in k[X]$ be a monic polynomial such that~$f(X) =
(X-x_1)\cdots(X-x_n)$ over~$L$. Let~$x$ be one of the zeros of~$f$. Show that
the theory shows~$\top \vdash \bigvee_{i=1}^n \sigma_{x,x_i}$.
\item Now furthermore assume that~$L$ is \emph{Galois over}~$k$ in the sense that
any element of~$L$ is the zero of a monic polynomial~$f \in k[X]$ which is
separable (that is,~$(f,f') = (1) \subseteq k[X]$) and splits over~$L$ into
linear factors. Show that the axioms expressing surjectivity can be dropped.
\item Show that the topological space of points of~$\Gal(L|k)$ is canonically
homeomorphic to the topological Galois group of~$L|k$.
\item Show that~$\Gal(L|k)$ is spatial in classical mathematics.
\end{enumerate}
\end{exercise}

\begin{exercise}{Coherent locales and coherent theories}%
\label{ex:coherent}%
An open~$U$ of a locale is \emph{compact} iff, whenever~$U
\preceq \bigvee_{i \in I} V_i$, there is a (Kuratowski-)finite subset~$I'
\subseteq I$ such that~$U \preceq \bigvee_{i \in I'} V_i$. A locale is
\emph{coherent} iff its compact opens are closed under finite meets and if
every open is a join of compact opens. A geometric formula is \emph{coherent} iff only
finite disjunctions occur in it. A geometric theory is \emph{coherent}
iff all of its axioms consist of coherent formulas.
\begin{enumerate}
\item Show that any coherent locale classifies some coherent theory.
\item Let~$\TT$ be a propositional coherent theory. Show that~$L(\TT)$ is
coherent.

@@

\item \emph{Deligne's completeness theorem} states that coherent locales (and
even coherent \emph{toposes}) are spatial. Show that Deligne's
result implies that propositional coherent theories
without any set-based models are inconsistent.
\end{enumerate}
\end{exercise}

\begin{exercise}{Subsheaves of the terminal sheaf}%
\label{ex:subsheaves-1}%
Let~$X$ be a locale.
\begin{enumerate}
\item Show that any subsheaf~$F \subseteq 1$ of the terminal sheaf on~$X$ gives rise
to an open of~$X$ by gluing all its sections.
\item Show that the construction of part~(a) yields a canonical bijection of
the subsheaves of~$1$ and the opens of~$X$.
\end{enumerate}
\end{exercise}

\begin{exercise}{A geometric interpretation of double negation}%
An open~$U$ of a locale is \emph{dense} iff for any open~$V$, $U \wedge V =
\bot$ implies~$U = \bot$. Let~$\varphi$ be a formula over a locale~$X$.
\begin{enumerate}
\item Show that there is a largest open,
denoted~``$\llbracket\varphi\rrbracket$'', such that~$U \models \varphi$.

@@

\item Show that~$X \models \neg\neg\varphi$ iff~$\llbracket\varphi\rrbracket$
is dense.
\end{enumerate}
\end{exercise}

@@

\begin{exercise}{The sheaf of real functions as a field}%
Let~$X$ be a topological space. Let~$\C$ be the sheaf of continuous real-valued
functions on~$X$.
\begin{enumerate}
\item Let~$U$ be an open of~$X$. Let~$f \in \C(U)$. Show using the result
of Exercise~\ref{ex:global-existence} that~$U \models (\exists g\?\C\_ fg =_\C
1)$ iff there is a function~$g \in \C(U)$ such that~$fg = 1$.
\item Show that~$\C$ is a field in that
$X \models \forall f\?\C\_
  \bigl(\neg(\exists g\?\C\_ fg = 1)\bigr) \Rightarrow f = 0$.
\item Give an example of space~$X$ such that it is not the case that~$\C$ is a
field in the stronger sense that
$\Sh(X) \models \forall f\?\C\_ (\exists g\?\C\_ fg = 1) \vee f = 0$.
\item Let~$X = \CC$ and let~$\O$ be the sheaf of holomorphic functions on~$X$.
Show that, in classical mathematics, the sheaf~$\O$ is \emph{discrete} in that
$\Sh(\CC) \models \forall f\?\O\_ \forall g\?\O\_ f = g \vee \neg(f = g)$.
\end{enumerate}
\end{exercise}

\begin{exercise}{On the failure of the fundamental theorem of algebra}%
\label{ex:fta}%
Let~$F$ be the sheaf of continuous complex-valued functions on~$\CC$. Give an
example of a function~$f : \CC \to \CC$ such that it is not the case that
\[ \Sh(X) \models \exists z\?F\_ z^2 = f. \]
{\scriptsize\emph{Note.} The sheaf~$F$ can be obtained by constructing, in the
universe of sheaves on~$\CC$, the set of pairs of Dedekind reals. Hence the
example demonstrates that the usual formulation of the fundamental theorem of
algebra does not admit a constructive proof. (However, there are replacements
available~\cite{richman:fta}, and the fundamental theorem does apply to the
Cauchy complex numbers~\cite{ruitenburg:roots} and under weak choice
principles~\cite{bridges-richman-schuster:wcc}).\par}
\end{exercise}

@@

The purpose of this section is to properly motivate the construction
leading to~$A^\sim$. Briefly, the story is as follows, with details given
below.
\begin{enumerate}
\item Given a ring~$A$, we set out to construct the \emph{free local ring}
over~$A$.
\item In the strict sense of the word, this endeavor will fail, but in a wider
sense the quest will succeed and yield the sheaf~$A^\sim$.
\item It is then a surprising observation that~$A^\sim$ is a field, even though
we only set out to construct a local ring.
\item The field property of~$A^\sim$ should be appropriately appreciated: While
there is a general machinery for free constructions of this kind, this
machinery is \emph{not} applicable for constructing free fields.
\end{enumerate}

@@

>>>>

\documentclass{ws-rv9x6}
\usepackage[onethmnum]{ws-rv-thm}
\usepackage{ws-rv-van}

% NOTE TO EDITORS AND LAYOUTERS: I require the following packages.
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{mathtools,stmaryrd,xspace,bussproofs}

\usepackage[english]{babel}

% NOTE TO EDITORS AND LAYOUTERS: This package increases the quality of the
% typesetting. I personally recommebd using it, but my contribution does not
% depend on it.
\usepackage[expansion=true,protrusion=true]{microtype}

\newtheorem{ex}{Exercise}
\let\Exercisefont\upshape
\def\Exerciseheadfont{\bfseries}
\newenvironment{exercise}[1]{
  \begin{ex}[#1]
}{\end{ex}}

\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red!80, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

@@

\begin{exercise}{The space of points of a locale}%
\label{ex:canonical-topology}%
Devise a canonical to\-po\-logy on the set of points of a locale.
\end{exercise}

\begin{exercise}{Embedding sober topological spaces into locales}%
Verify the statements made in Remark~\ref{rem:sober-embed}.
\end{exercise}

\begin{exercise}{Hausdorff spaces are sober}%
\label{ex:hausdorff-sober}%
A \emph{meet-irreducible open subset} of a topological space~$Y$ is an
open subset~$W \subseteq Y$ such that~$(U \cap V \subseteq W) \Rightarrow (U
\subseteq W) \vee (V \subseteq W)$ for all open subsets~$U,V \subseteq Y$ and
such that~$Y \not\subseteq W$.
\begin{alphlist}[(c)]
\item Assuming the law of excluded middle, show that the completely prime
filters of a topological space are in canonical one-to-one correspondence with
the meet-irreducible open subsets.\smallskip

@@

\item Assuming the law of excluded middle, show that Hausdorff topological
spaces validate the condition stated in part~(b) and hence are sober.
\end{alphlist}
{\scriptsize\emph{Note.} Without the law of excluded middle, Hausdorff spaces
and even metric spaces may fail to be sober. In fact, already the topological
space of the rational numbers can fail to be
sober~\cite[Example~8.14(iv)]{fourman-scott:sheaves-and-logic}.\par}
\end{exercise}

\begin{exercise}{Complete metric spaces are sober}
A \emph{metric space} is a set~$Y$ together with a relation~$d \subseteq Y
\times Y \times \QQ_{\geq0}$, written~``$d(x,y) \leq q$'', such that the
following axioms are satisfied.
\begin{align*}
  d(x,y) \leq q \wedge q \leq r &\Longrightarrow d(x,y) \leq r \\
  (\forall r \in \QQ_{> q}\_ d(x,y) \leq r) &\Longrightarrow d(x,y) \leq q \\
  &\mathrel{\phantom{\Longrightarrow}} \exists q \in \QQ_{\geq0}\_ d(x,y) \leq q \\
  &\mathrel{\phantom{\Longrightarrow}} d(x,x) \leq 0 \\
  d(x,y) \leq 0 &\Longrightarrow x = y \\
  d(x,y) \leq q &\Longrightarrow d(y,x) \leq q \\
  d(x,y) \leq q \wedge d(y,z) \leq r &\Longrightarrow d(x,z) \leq q+r
\end{align*}
A \emph{Cauchy process} (or \emph{multi-valued modulated Cauchy sequence}) in a metric
space~$Y$ is a map~$\alpha : \QQ_{>0} \to P(Y)$ such that all sets~$\alpha(q)$
are inhabited and such that~$d(x,x') \leq q + q'$ for all~$q,q' \in \QQ_{>0}$
and~$x \in \alpha(q), x' \in \alpha(q')$. Such a process \emph{converges
to}~$x_0 \in Y$ iff~$d(x,x_0) \leq q$ for all~$q \in \QQ_{>0}$
and~$x \in \alpha(q)$. A metric space~$Y$ is \emph{complete} iff every Cauchy
process in~$Y$ converges.

Verify, without using the law of excluded middle, that any complete metric
space is sober.
\end{exercise}

@@

\begin{exercise}{The covariant approach to locales}%
Let~$A$ be a frame.
\begin{alphlist}[(c)]
\item Show that~$A$ contains arbitrary (set-indexed) meets (though they are not
required to be preserved by frame homomorphisms), by the impredicative
construction
\[ \bigwedge_{i \in I} U_i = \bigvee\{ \text{$V \preceq U_i$ for all~$i \in I$}
\}. \]
\item What are the arbitrary meets in the frame of open subsets of a
topological space?
\item Show that for any frame homomorphism~$\alpha : A \to A'$ there exists a
monotone map~$\beta : A' \to A$ such that~\emph{$\beta$ is right adjoint
to~$\alpha$}, that is such that
\[ \alpha(U) \preceq V \quad\text{iff}\quad U \preceq \beta(V) \]
for all~$U \in A$ and~$V \in A'$.
\end{alphlist}
{\scriptsize\emph{Note.} Based on the observation in part~(c), a
\emph{covariant approach} to locales can be developed, in which locale maps~$X
\to X'$ are defined to be certain kinds of maps~$\O(X) \to
\O(X')$~\cite{picado-pultr:covariant}.\par}
\end{exercise}

@@

\begin{exercise}{Characterizing spatial locales}%
Let~$X$ be a locale. Show that the following conditions are equivalent.
\begin{alphlist}[(c)]
\item For any opens~$U, V \in \O(X)$: If~$x \inplus U \Rightarrow y \inplus V$
for all points~$x$ of~$X$, then~$U \preceq V$.
\item For any opens~$U, V \in \O(X)$: If~$x \inplus U \Leftrightarrow y \inplus V$
for all points~$x$ of~$X$, then~$U = V$.
\item There is a topological space~$Y$ and a surjective morphism~$L(Y) \to X$
of locales. (A morphism of locales is \emph{surjective} iff its underlying
frame homomorphism is injective.)
\item The canonical morphism~$L(\Pt(X)) \to X$ is an isomorphism.
\end{alphlist}
\end{exercise}

\begin{exercise}{The frame of radical ideals}%
\label{ex:frame-of-radical-ideals}%
Let~$A$ be a ring. We order the set~$\Idl(A)$ of ideals of~$A$ by inclusion.
(See Section~\ref{sect:algebraic-preliminaries} for some algebraic
preliminaries.)
\begin{alphlist}[(b)]
\item Verify that~$\Idl(A)$ has arbitrary meets and joins, given by
intersection and sum of ideals, but that it is in general not a
frame.\smallskip

{\scriptsize\emph{Hint.} The distributive law fails in the polynomial ring~$\QQ[X,Y]$.\par}
\item Show that the sub-poset~$\Rad(A)$ of radical ideals is a frame, with
the same finite meets but with joins given by the radical of the join computed
in~$\Idl(A)$.
\item\label{item:join-top} Let~$\varphi$ be a truth value. Let~$\{\top\,|\,\varphi\}$ be the
subsingleton subset of~$\Rad(A)$ which contains the top element~$\sqrt{(1)}$
iff~$\varphi$. Verify that
\[ \bigvee\{\top\,|\,\varphi\} = \{ f \in A \,|\, \text{$f$ is nilpotent
or~$\varphi$} \} \]
in~$\Rad(A)$.
\end{alphlist}
\end{exercise}

\begin{exercise}{Coproducts of locales}%
\begin{alphlist}[(c)]
\item Show that the category of frames has small products.
\item Deduce that the category of locales has small coproducts.
\item Let~$M$ be a set and endow it with the discrete topology. Show
that~$L(M)$ is the coproduct~$\coprod_{x \in M} \pt$.
\end{alphlist}
{\scriptsize\emph{Note.} The category of locales also has small products;
Exercise~\ref{ex:products-of-locales} is devoted to this fact.\par}
\end{exercise}

@@

\begin{exercise}{Maps into discrete spaces}%
Let~$X$ be a locale. Let~$A$ be a set and endow it with the discrete topology. Show
that morphisms~$X \to L(A)$ are in canonical correspondence with
families~$(U_a)_{a \in A}$ of opens of~$X$ such that~$\top = \bigvee_{a \in A}
U_a$ and~$U_a \wedge U_b \preceq \bigvee\{ \top \,|\, a = b \}$.
\end{exercise}

@@

\begin{exercise}{Discrete locales as classifying locales}%
\begin{alphlist}[(b)]
\item Find a theory which is classified by the two-point locale (the locale
associated to the discrete two-point topological space).
\item Let~$M$ be a set and endow it with the discrete topology. Find a theory
which is classified by~$L(M)$.
\end{alphlist}
\end{exercise}

@@

\begin{exercise}{More on exotic locales}%
\label{ex:exotic-locales}%
\begin{alphlist}[(d)]
\item Define the locale of injections~$\RR \hookrightarrow \NN$.
\item Define the locale of partial surjections~$\NN \twoheadrightarrow \NN^\NN$.\smallskip

{\scriptsize\emph{Note.} In Russian constructivism, as for instance reigning in
the ef{}fective topos~\cite{hyland:effective-topos,phoa:effective,bauer:c2c}, the Baire space~$\NN^\NN$ is subcountable.
However, in this setting any function~$\NN \to \NN$ is computable.
The universe of sheaves over this locale was the first example of a setting in
which the Baire space is subcountable and still any external number-theoretic
function can be imported~\cite[Section~4.3]{scedrov:forcing}.\par}
\item Define the locale of ``open covers of~$[0,1]$ by intervals with rational
endpoints such that the union of any finite number of those intervals has total
Lebesgue measure less than~$\frac{1}{2}$''.\smallskip

{\scriptsize\emph{Note.} In the universe of sheaves over this locale, the set
of unit interval Dedekind real numbers fails to be (Heine--Borel) compact and
the localic real line fails to be
spatial~\cite[Example~D4.7.13]{johnstone:elephant}.\par}
\item An open~$U$ of a locale~$X$ is \emph{complemented} iff there is an
open~$V$ such that~$U \vee V = \top$ and~$U \wedge V = \bot$. A locale is
\emph{zero-dimensional} iff any open is a join of complemented opens. Assuming
that any two reals are equal or not, show that the locale of surjections~$\NN
\twoheadrightarrow \RR$ is zero-dimensional.
\end{alphlist}
\end{exercise}

\begin{exercise}{Models as points}%
Let~$\TT$ be a propositional geometric theory. Verify that the points
of~$L(\TT)$ are in canonical bijection with the set-based models of~$\TT$.
\end{exercise}

@@

\begin{exercise}{Spatiality of the localic reals}%
\label{ex:spatiality-reals}%
\newcommand{\fa}{\ensuremath{(\!\cdot}}%
\newcommand{\fb}{\ensuremath{\cdot\!)}}%
Let~$\RR$ be the localic real line, defined as the classifying locale of the
theory~$\TT$ of Dedekind cuts as as introduced on
page~\pageref{par:dedekind-cuts}. For rational numbers~$x,y$, let~``$\fa x,y
\fb$'' denote the formula~$\alpha_x \wedge \beta_y$. Let~$(x,y)$ denote the
open interval~$\{ \xi \in \Pt(\RR) \,|\, x < \xi < y \}$ in the set of Dedekind reals.
\begin{alphlist}[(f)]
\item Extend the notation to allow for~$x,y=\pm\infty$.
\item Show that the canonical topology on~$\Pt(\RR)$, defined in
Exercise~\ref{ex:canonical-topology}, coincides with the usual Euclidean
topology.
\item Show that any open of~$\RR$ can be written in the form~$\bigvee_i
[\fa x_i,y_i \fb]$ with~$x_i,y_i \in \QQ \cup \{\pm\infty\}$.
\item Let rational numbers be given such that~$(x,y) \subseteq \bigcup_i
(x_i,y_i)$ in~$\Pt(\RR)$. Assume that this open covering has a Lebesgue number.
Show that~$\TT$ proves~$\fa x,y \fb \vdash \bigvee_i \fa x_i,y_i \fb$.
\item Assume that closed intervals in the set of Dedekind reals are
(Heine--Borel) compact. Then redo part~(d), but without the assumption on the
existence of a Lebesgue number.
\item Under the same assumption as in part~(e), show that the localic real line
is spatial.
\end{alphlist}
{\scriptsize\emph{Note.} Exercise~\ref{ex:exotic-locales} lists settings in
which the assumption in part~(e) is not satisfied.\par}
\end{exercise}

\begin{exercise}{The localic Galois group of a field extension}%
Let~$L|k$ be a field extension. The \emph{localic Galois group}~$\Gal(L|k)$
is the classifying locale of the theory of automorphisms~$L \to L$ which
fix~$k$, as described on page~\pageref{par:theory-automorphisms}.
\begin{alphlist}[(g)]
\item Explain how to obtain a morphism~$\Gal(L|k) \to \Gal(L|k)$ which on the
level of points should send an automorphism~$\sigma$ to its inverse~$\sigma^{-1}$.
\item Explain how to obtain a morphism~$\Gal(L|k) \times \Gal(L|k) \to \Gal(L|k)$ which on the
level of points should send a pair~$\langle \sigma, \tau \rangle$ to~$\sigma \circ \tau$.
\item Assume from now on that~$L$ is a \emph{geometric field} in the sense that any
element is zero or invertible. Show that the axiom expressing injectivity can
be dropped without changing the set of provable sequents.
\item Let~$f \in k[X]$ be a monic polynomial such that~$f(X) =
(X-x_1)\cdots(X-x_n)$ over~$L$. Let~$x$ be one of the zeros of~$f$. Show that
the theory shows~$\top \vdash \bigvee_{i=1}^n \sigma_{x,x_i}$.
\item Now furthermore assume that~$L$ is \emph{Galois over}~$k$ in the sense that
any element of~$L$ is the zero of a monic polynomial~$f \in k[X]$ which is
separable (that is,~$(f,f') = (1) \subseteq k[X]$) and splits over~$L$ into
linear factors. Show that the axioms expressing surjectivity can be dropped.
\item Show that the topological space of points of~$\Gal(L|k)$ is canonically
homeomorphic to the topological Galois group of~$L|k$.
\item Show that~$\Gal(L|k)$ is spatial in classical mathematics.
\end{alphlist}
\end{exercise}

\begin{exercise}{Coherent locales and coherent theories}%
\label{ex:coherent}%
An open~$U$ of a locale is \emph{compact} iff, whenever~$U
\preceq \bigvee_{i \in I} V_i$, there is a (Kuratowski-)finite subset~$I'
\subseteq I$ such that~$U \preceq \bigvee_{i \in I'} V_i$. A locale is
\emph{coherent} iff its compact opens are closed under finite meets and if
every open is a join of compact opens. A geometric formula is \emph{coherent} iff only
finite disjunctions occur in it. A geometric theory is \emph{coherent}
iff all of its axioms consist of coherent formulas.
\begin{alphlist}[(d)]
\item Show that any coherent locale classifies some coherent theory.
\item Let~$\TT$ be a propositional coherent theory. Show that~$L(\TT)$ is
coherent.\smallskip

@@

\item \emph{Deligne's completeness theorem} states that coherent locales (and
even coherent \emph{toposes}) are spatial. Show that Deligne's
result implies that propositional coherent theories
without any set-based models are inconsistent.
\end{alphlist}
\end{exercise}

\begin{exercise}{Subsheaves of the terminal sheaf}%
\label{ex:subsheaves-1}%
Let~$X$ be a locale.
\begin{alphlist}[(b)]
\item Show that any subsheaf~$F \subseteq 1$ of the terminal sheaf on~$X$ gives rise
to an open of~$X$ by gluing all its sections.
\item Show that the construction of part~(a) yields a canonical bijection of
the subsheaves of~$1$ and the opens of~$X$.
\end{alphlist}
\end{exercise}

\begin{exercise}{A geometric interpretation of double negation}%
An open~$U$ of a locale is \emph{dense} iff for any open~$V$, $U \wedge V =
\bot$ implies~$U = \bot$. Let~$\varphi$ be a formula over a locale~$X$.
\begin{alphlist}[(b)]
\item Show that there is a largest open,
denoted~``$\llbracket\varphi\rrbracket$'', such that~$U \models \varphi$.\smallskip

@@

\item Show that~$X \models \neg\neg\varphi$ iff~$\llbracket\varphi\rrbracket$
is dense.
\end{alphlist}
\end{exercise}

@@

\begin{exercise}{The sheaf of real functions as a field}%
Let~$X$ be a topological space. Let~$\C$ be the sheaf of continuous real-valued
functions on~$X$.
\begin{alphlist}[(d)]
\item Let~$U$ be an open of~$X$. Let~$f \in \C(U)$. Show using the result
of Exercise~\ref{ex:global-existence} that~$U \models (\exists g\?\C\_ fg =_\C
1)$ iff there is a function~$g \in \C(U)$ such that~$fg = 1$.
\item Show that~$\C$ is a field in that
$X \models \forall f\?\C\_
  \bigl(\neg(\exists g\?\C\_ fg = 1)\bigr) \Rightarrow f = 0$.
\item Give an example of space~$X$ such that it is not the case that~$\C$ is a
field in the stronger sense that
$\Sh(X) \models \forall f\?\C\_ (\exists g\?\C\_ fg = 1) \vee f = 0$.
\item Let~$X = \CC$ and let~$\O$ be the sheaf of holomorphic functions on~$X$.
Show that, in classical mathematics, the sheaf~$\O$ is \emph{discrete} in that
$\Sh(\CC) \models \forall f\?\O\_ \forall g\?\O\_ f = g \vee \neg(f = g)$.
\end{alphlist}
\end{exercise}

\begin{exercise}{The failure of the fundamental theorem of algebra}%
\label{ex:fta}%
Let~$F$ be the sheaf of continuous complex-valued functions on~$\CC$. Give an
example of a function~$f : \CC \to \CC$ such that it is not the case that
\[ \Sh(\CC) \models \exists z\?F\_ z^2 = f. \]
{\scriptsize\emph{Note.} The sheaf~$F$ can be obtained by constructing, in the
universe of sheaves on~$\CC$, the set of pairs of Dedekind reals. Hence the
example demonstrates that the usual formulation of the fundamental theorem of
algebra does not admit a constructive proof. (However, there are replacements
available~\cite{richman:fta}, and the fundamental theorem does apply to the
Cauchy complex numbers~\cite{ruitenburg:roots} and under weak choice
principles~\cite{bridges-richman-schuster:wcc}).\par}
\end{exercise}

@@

The purpose of this section is to properly motivate the construction
leading to~$A^\sim$. Briefly, the story is as follows, with details given
below.
\begin{enumerate}
\item Given a ring~$A$, we set out to construct the \emph{free local ring}
over~$A$.
\item In the strict sense of the word, this endeavor will fail, but in a wider
sense the quest will succeed and yield the sheaf~$A^\sim$.
\item It is then a surprising observation that~$A^\sim$ is a field, even though
we only set out to construct a local ring.
\item The field property of~$A^\sim$ should be appropriately appreciated: While
there is a general machinery for free constructions of this generalized kind, this
machinery is \emph{not} applicable for constructing free fields.
\end{enumerate}

@@

\begin{remark}There is general topos-theoretic machinery for generalized free
constructions of this kind~\cite{coste:sheaf-representation,cole:spectra}, and
this machinery can be used to construct several kinds of spectra, including the
ordinary spectrum of a ring. However, this machinery does have limitations, and
in particular it cannot be used to construct free fields.
\end{remark}
